---
title: "Analysis Script"
author: "CV_CS"
date: "30 12 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis Script Bias in AI

Load relevant packages:
```{r}
library(readxl)
library(tidyverse)
```




## Read in file exported from REDCap
```{r}
data <- read_excel("BiasUndKI_DATA_2021-12-09_1601.xlsx")
View(data)
```


## Prepare Analysis Data 

divide dataset depending on the chosen language

### German
```{r}
german <- data[data$language == 1, ]
nrow(german)
```

Remove NA Fields in language (no data was submitted by user)
```{r}
german <- german %>% drop_na(language)
nrow(german)
```
#### Delete Columns from English Speaking Participants
German Questionnaire starts with the following column:"fragebogen_fairness_und_ki_timestamp"
English Questionnaire starts with the following column:"questionnaire_fairness_and_ai_timestamp" 
Detect Column number

```{r}
which(colnames(german) == "fragebogen_fairness_und_ki_timestamp")
which(colnames(german) == "questionnaire_fairness_and_ai_timestamp")
```

Delete these columns
```{r}
german[ , 141:275] <- list(NULL)
```
```{r}
ncol(german)
```


### English
```{r}
En <- data[data$language == 2, ]
nrow(En)
```

Remove NA Fields in language (no data was submitte by user)
```{r}
En <- En %>% drop_na(language)
nrow(En)
```

#### Delete Columns from German Participants
German Questionnaire starts with the following column:"fragebogen_fairness_und_ki_timestamp"
English Questionnaire starts with the following column:"questionnaire_fairness_and_ai_timestamp"  
Detect Column number
```{r}
which(colnames(En) == "fragebogen_fairness_und_ki_timestamp")
which(colnames(En) == "questionnaire_fairness_and_ai_timestamp")
```

Delete these columns
```{r}
En[ , 6:140] <- list(NULL)
```
```{r}
ncol(En)
```
### Prepare Data for Merging 
Both dataframes need same column names
Export into excel and copy column names from german into En, reimport data
(Did not find this code in R)


```{r}
library(writexl)
write_xlsx(german, 'german.xlsx')
write_xlsx(En, 'En.xlsx')
```

```{r}
german_merge <- read_excel("german_merge.xlsx")
En_merge <- read_excel("En_merge.xlsx")
View(En_merge)
View(german_merge)
```


## Merge Datasets
```{r}
merged <- rbind(german_merge, En_merge)
```

***
## Analysis

```{r}
nrow(merged)
ncol(merged)
```
Delete NA only columns
```{r}
merged <- merged[ , colSums(is.na(merged)) < nrow(merged)]
```

### Total Participants
```{r}
nrow(merged)
ncol(merged)
```

### questionnaire 1 complete?
```{r}
complete1 <- merged[merged$fragebogen_fairness_und_ki_complete == 2, ]
nrow(german)
```


### KI Entwickler?
1-3
```{r}
is.factor(complete1$ki_entwicklung)
```
```{r}
complete1$ki_entwicklung = as.factor(complete1$ki_entwicklung)
```


```{r}
developer <- complete1 %>% 
  group_by(ki_entwicklung) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>% 
  arrange(desc(N))

#Rename Variables
developer$ki_entwicklung <- factor(developer$ki_entwicklung,levels = c(1,2,3,4,5), labels = c("Yes","No, but I have worked on AI-projects in the past", "No, but I am planning on it", "No, never", "Not specified"))

knitr::kable(developer)
```

```{r}
# define theme and colors for barplots
bar_theme <-   theme(rect = element_blank(),
        axis.ticks.y = element_line(color = "grey40"),
        panel.grid.major.x = element_line(color = "grey90"))
bar_color <-  "#1b3051"
bar_width <- 0.6
```

```{r}
ggplot(developer) + geom_col(aes(reorder(ki_entwicklung, N, sum), N),
                              fill = bar_color,
                              width = bar_width) +
  labs(x = "", y = "Sind Sie derzeit an einer KI-Entwicklung beteiligt?") +
  scale_y_continuous(breaks = seq(0,200,50)) +
  coord_flip() +
  bar_theme
```
Data of developers only
```{r}
nrow(complete1)
```
```{r}
AIcomplete1 <- complete1[complete1$ki_entwicklung == 1 | complete1$ki_entwicklung == 2 | complete1$ki_entwicklung == 3, ]
nrow(AIcomplete1)
```

### questionnaire 2 complete? 




























##Save Code for further Analysis

Drop Variables where All Values are Missing (German Participants)
```{r}
test <- En[ , colSums(is.na(En)) < nrow(En)]
```


