---
title: "Data Cleaning for Analysis"
author: "Vorisek_CN"
date: "17 01 2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare Data Cleaningda
*Load relevant libraries*
```{r}
library(readxl)
library(tidyverse)
library(writexl)
library(knitr)
```

*Bar themes*
```{r}
# define theme and colors for barplots
bar_theme <-   theme(rect = element_blank(),
        axis.ticks.y = element_line(color = "grey40"),
        panel.grid.major.x = element_line(color = "grey90"))
bar_color <-  "#1b3051"
bar_width <- 0.6
```

*Read in file created in "Data_Preparation.Rmd"*
```{r}
merged <- read_excel("merged.xlsx")
View(merged)
```

**Entire Data**
```{r}
nrow(merged)
ncol(merged)
```


# Delete Incomplete Datasets of AI Developers
## Questionnaire 1 completed

```{r}
complete1 <- merged[merged$fragebogen_fairness_und_ki_complete == 2, ]
nrow(complete1)
```
## Who completed sheet 1?
```{r}
is.factor(complete1$ki_entwicklung)
```
```{r}
complete1$ki_entwicklung = as.factor(complete1$ki_entwicklung)
```

```{r}
developer <- complete1 %>% 
  group_by(ki_entwicklung) %>% 
  summarize(N = n()) %>% 
  mutate(Percent = N / sum(N) * 100,
         Percent = round(Percent, 1)) %>% 
  arrange(desc(N))

#Rename Variables
developer$ki_entwicklung <- factor(developer$ki_entwicklung,levels = c(1,2,3,4,5), labels = c("Yes","No, but I have worked on AI-projects in the past", "No, but I am planning on it", "No, never", "Not specified"))

knitr::kable(developer)
```

```{r}
ggplot(developer) + geom_col(aes(reorder(ki_entwicklung, N, sum), N),
                              fill = bar_color,
                              width = bar_width) +
  labs(x = "", y = "Sind Sie derzeit an einer KI-Entwicklung beteiligt?") +
  scale_y_continuous(breaks = seq(0,200,50)) +
  coord_flip() +
  bar_theme
```


## Data of AI Developers only
```{r}
AIcomplete1 <- complete1[complete1$ki_entwicklung == 1 | complete1$ki_entwicklung == 2 | complete1$ki_entwicklung == 3, ]
nrow(AIcomplete1)
```
## Questionnaire 2 completed
```{r}
AIcomplete2 <- AIcomplete1[AIcomplete1$fragebogen_fairness_und_ki_2_complete == 2, ]
nrow(AIcomplete2)
```
## response rate
```{r}
199/334*100
```

# Identify additional Incomplete IDs
*"Show only fields" are not included to identify missing data, evaluation in a second step*

**Demographics**

*Prepare Data*
```{r}
AIcomplete2$arbeitsland = as.factor(AIcomplete2$arbeitsland)
AIcomplete2$geschlecht = as.factor(AIcomplete2$geschlecht)
AIcomplete2$sektor = as.factor(AIcomplete2$sektor)
AIcomplete2$quelle = as.factor(AIcomplete2$quelle)
AIcomplete2$alter = as.numeric(AIcomplete2$alter)
```


## How did this questionnaire reach you?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(quelle)
nrow(AIcomplete2)
```

## Which country do you work in?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(arbeitsland)
nrow(AIcomplete2)
```

## What is your gender?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(geschlecht)
nrow(AIcomplete2)
```

## what is your age?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(alter)
nrow(AIcomplete2)
```

## In which area do you work in?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(sektor)
nrow(AIcomplete2)
```


## At what stage are you or have you been developing AI-projects?
```{r}
which(colnames(AIcomplete2) == "ki_entwicklungsphase___1")
which(colnames(AIcomplete2) == "ki_entwicklungsphase___8")
```
*create dataset*
```{r}
data_ID<-AIcomplete2 %>% select(1, 19:26)
view(data_ID)
ncol(data_ID)
```
**identify incomplete IDs**
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:9) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```

*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <-AIcomplete2[!(AIcomplete2$id==136 | AIcomplete2$id==439 | AIcomplete2$id==625),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```


## In which medical specialty are you developing AI solutions?
```{r}
which(colnames(AIcomplete2) == "med_fachbereich___1")
which(colnames(AIcomplete2) == "med_fachbereich___37")
```
*create dataset*
```{r}
data_ID<-AIcomplete2 %>% select(1, 28:64)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:38) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==36 | AIcomplete2$id==143 | AIcomplete2$id==278 | AIcomplete2$id==567 | AIcomplete2$id==609),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```




## In which of the following areas of AI are you developing?
```{r}
which(colnames(AIcomplete2) == "ki_art___1")
which(colnames(AIcomplete2) == "ki_art___4")
```
*create dataset*
```{r}
data_ID<-AIcomplete2 %>% select(1,  66:69)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:5) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==316  | AIcomplete2$id==366),]
nrow(AIcomplete2)
rm(data_ID)
rm(data_ID)
```

## In which of the following areas of AI are you developing?

```{r}
which(colnames(AIcomplete2) == "ki_bereich___1")
which(colnames(AIcomplete2) == "ki_bereich___7")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1,  81:87)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:8) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==231  | AIcomplete2$id==351  | AIcomplete2$id==357 | AIcomplete2$id==560),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```


*** 
**Questions of Interest**
## How familiar are you with biases in AI?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(bias_ki)
nrow(AIcomplete2)
```

## Where do you think biases in AI can occur?

```{r}
which(colnames(AIcomplete2) == "bias_feld___1")
which(colnames(AIcomplete2) == "bias_feld___6")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1,  90:95)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:7) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==247   | AIcomplete2$id==257   | AIcomplete2$id==269  | AIcomplete2$id==290  | AIcomplete2$id==382    | AIcomplete2$id==422   | AIcomplete2$id==624),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```


## What data do you use to train AI algorithms?
```{r}
which(colnames(AIcomplete2) == "ki_trainingsdaten___1")
which(colnames(AIcomplete2) == "ki_trainingsdaten___4")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1, 97:100)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:5) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==214    | AIcomplete2$id==330    | AIcomplete2$id==587),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```


## What is the origin of your data?
```{r}
which(colnames(AIcomplete2) == "ki_daten_quelle___1")
which(colnames(AIcomplete2) == "ki_daten_quelle___6")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1, 101:106)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:7) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==152),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```



## Do you work with national or international data?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(nat_internat_daten)
nrow(AIcomplete2)
```

## Do you think that using standardized data (internationalsemantic and syntactic standards such as HL7 FHIR, SNOMEDCT) when training AI algorithms can prevent bias?
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(standard_daten)
nrow(AIcomplete2)
```


## Do you know any of the following preventive measures to avoidbias in AI applications?
```{r}
which(colnames(AIcomplete2) == "massnahmen_bias___1")
which(colnames(AIcomplete2) == "massnahmen_bias___6")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1, 110:115)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:7) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2[!(AIcomplete2$id==529      | AIcomplete2$id==28),]
nrow(AIcomplete2)
rm(data_ID)
rm(incomplete)
```

## What sociodemographic data would you collect to prevent biases in AI?
```{r}
which(colnames(AIcomplete2) == "soziodem_daten___1")
which(colnames(AIcomplete2) == "soziodem_daten___7")
```
*create dataset*
```{r}
data_ID <-AIcomplete2 %>% select(1, 117:123)
view(data_ID)
ncol(data_ID)
```
*identify incomplete IDs*
```{r}
data_ID <- data_ID %>%
    mutate(nonresp = select(., 2:8) %>% rowSums(na.rm = TRUE))
incomplete <- data_ID[data_ID$nonresp == 0, ]
incomplete$id
```
*drop non resp*
```{r}
rm(data_ID)
rm(incomplete)
```

## How would you rate the level of fairness of AI in your AIdevelopment? (Are specifi c groups of people treated /considereddiff erently from others?)

```{r}
nrow(AIcomplete2)
AIcomplete2 <- AIcomplete2 %>% drop_na(fairness_ki)
nrow(AIcomplete2)
```

# Create File for Analysis
```{r}
write_xlsx(AIcomplete2, 'AnalysisFile_Bias_AI.xlsx')
```






















































