---
title: "Data Preparation from RedCap"
author: "Vorisek_CN"
date: "16 1 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare Data from RedCap for Analysis
## Load relevant packages:
```{r}
library(readxl)
library(tidyverse)
library(writexl)
library(knitr)
```

## Read in file exported from REDCap
```{r}
data <- read_excel("BiasUndKI_DATA_2021-12-09_1601.xlsx")
View(data)
```

## Divide dataset depending on the chosen language
*otherwise double entry due to longitudinal data*

### German Speaking Participants
```{r}
german <- data[data$language == 1, ]
nrow(german)
```
*Remove NA Fields in language (no data was submitted by user)*
```{r}
german <- german %>% drop_na(language)
nrow(german)
```
#### Delete Columns from English Speaking Participants
*German Questionnaire starts with the following column:"fragebogen_fairness_und_ki_timestamp"*
*English Questionnaire starts with the following column:"questionnaire_fairness_and_ai_timestamp"* 
**Detect Column number**
```{r}
which(colnames(german) == "fragebogen_fairness_und_ki_timestamp")
which(colnames(german) == "questionnaire_fairness_and_ai_timestamp")
```
**Delete these columns**
```{r}
german[ , 141:275] <- list(NULL)
ncol(german)
```

### English Speaking Participants
```{r}
En <- data[data$language == 2, ]
nrow(En)
```
*Remove NA Fields in language (no data was submitted by user)*
```{r}
En <- En %>% drop_na(language)
nrow(En)
```
#### Delete Columns from English Speaking Participants
*German Questionnaire starts with the following column:"fragebogen_fairness_und_ki_timestamp"*
*English Questionnaire starts with the following column:"questionnaire_fairness_and_ai_timestamp"* 
**Detect Column number**
```{r}
which(colnames(En) == "fragebogen_fairness_und_ki_timestamp")
which(colnames(En) == "questionnaire_fairness_and_ai_timestamp")
```
**Delete these columns**
```{r}
En[ , 6:140] <- list(NULL)
ncol(En)
```


# Merging the German and English Datasets
## Prepare Data for Merging 
*Both dataframes need same column names*
```{r}
write_xlsx(german, 'german.xlsx')
write_xlsx(En, 'En.xlsx')
```

**Export into excel and copy column names from german into En, reimport data**
*(Did not find this code in R)*
```{r}
german_merge <- read_excel("german_merge.xlsx")
En_merge <- read_excel("En_merge.xlsx")
View(En_merge)
View(german_merge)
```

## Merge Datasets
```{r}
merged <- rbind(german_merge, En_merge)
```


# Prepare Analysis
## Identify relevant data
*Do columns add up?*
```{r}
nrow(En)
nrow(german)
nrow(merged)
ncol(merged)
```
*Delete NA only columns*
```{r}
merged <- merged[ , colSums(is.na(merged)) < nrow(merged)]
```

# Write File for Analysis
```{r}
write_xlsx(merged, 'merged.xlsx')
```
